---
title: "R Notebook"
output: html_notebook
---

1.3.2 Industrial and Labor Relations
In 1947, the United States Congress passed the Taft-Hartley Amendments to the
Wagner Act. The original Wagner Act had permitted the unions to use a Closed
Shop Contracf unless prohibited by state law. The Taft-Hartley Amendments
made the use of Closed Shop Contract illegal and gave individual states the right
to prohibit union shops6 as well. These right-to-work laws have caused a wave
of concern throughout the labor movement. A question of interest here is: What
are the effects of these laws on the cost of living for a four-person family living
on an intermediate budget in the United States? To answer this question a data set
consisting of 38 geographic locations has been assembled from various sources.
The variables used are defined in Table 1.2. The Right-To-Work Laws data are
given in Table 1.3 and can also be found at the book's Website.

Variable
COL     Cost of living for a four-person family
PD      Population density (person per square mile)
URate   State unionization rate in 1978
Pop     Population in 1975
Taxes   Property taxes in 1972
Income  Per capita income in 1974
RTWL    Indicator variable (1 if there are right-to-work laws in the state and 0                  otherwise)

```{r}

library(dplyr)
library(scatterplot3d)
library(psych)
library(ggplot2)
library(car)
library(gclus)
library(GGally)
library(fields)
```


```{r}
data <- read.delim("P005.txt")
rtwl_data<-data[,-1]
```


```{r}
plot(rtwl_data)
```
```{r}
model_null <- lm( COL~1, data = rtwl_data)
model_full <- lm( COL~PD*URate*Pop*Taxes*Income*RTWL, data=rtwl_data)


```

```{r}
step(model1, scope=list(lower=model_null, 
                        upper=model_full),
     direction="forward")
```
```{r}
fit <- lm(formula = COL ~ RTWL + Pop + URate + 
            Income + PD + RTWL:URate + 
            RTWL:Income + URate:Income, 
          data = rtwl_data)
```

```{r}
summary(fit)
```

Normal prob plot
```{r}
# qq plot for studentized resid
qqPlot(fit, main="QQ Plot")
```

Pred. vs StRes.
```{r}
residualPlots(fit)
```






```{r}
# Assessing Outliers
outlierTest(fit) # Bonferonni p-value for most extreme obs
qqPlot(fit, main="QQ Plot") #qq plot for studentized resid 
leveragePlots(fit) # leverage plots
```
```{r}
# Influential Observations
# added variable plots 
avPlots(fit)
# Cook's D plot
# identify D values > 4/(n-k-1) 
cutoff <- 4/((nrow(mtcars)-length(fit$coefficients)-2)) 
plot(fit, which=4, cook.levels=cutoff)
# Influence Plot 
influencePlot(fit,	id="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )
```

```{r}
# Normality of Residuals
# qq plot for studentized resid
qqPlot(fit, main="QQ Plot")
# distribution of studentized residuals
library(MASS)
sresid <- studres(fit) 
hist(sresid, freq=FALSE, 
   main="Distribution of Studentized Residuals")
xfit<-seq(min(sresid),max(sresid),length=40) 
yfit<-dnorm(xfit) 
lines(xfit, yfit)
```

```{r}
# Evaluate homoscedasticity
# non-constant error variance test
ncvTest(fit)
# plot studentized residuals vs. fitted values 
spreadLevelPlot(fit)
```

```{r}
# Evaluate Collinearity
vif(fit) # variance inflation factors 
sqrt(vif(fit)) > 2 # problem?
```


```{r}
# Test for Autocorrelated Errors
durbinWatsonTest(fit)
```

```{r}
# Global test of model assumptions
library(gvlma)
gvmodel <- gvlma(fit) 
summary(gvmodel)
```

```{r}

```

```{r}

```

```{r}

```

```{r}
 #Making a Basic scatter plot matrix
 
 scatterplotMatrix(~COL+PD+URate+Pop+Taxes+Income+RTWL ,data=rtwl_data, smooth = FALSE, ellipse=
 FALSE, main="Simple Scatterplot Matrix")

```


```{r}
dta.r <- abs(cor(rtwl_data)) # get correlations
dta.col <- dmat.color(dta.r) # get colors
# reorder variables so those with highest correlation
# are closest to the diagonal
dta.o <- order.single(dta.r) 
cpairs(rtwl_data, dta.o, panel.colors=dta.col, gap=.5,
main="Variables Ordered and Colored by Correlation" )

dta.r
```


attach(rtwl_data) 
s3d <-scatterplot3d(COL,Income,RTWL, pch=16, highlight.3d=TRUE,
  type="h", main="3D Scatterplot")
fit <- lm(COL ~ Income+ RTWL, data = rtwl_data) 
s3d$plane3d(fit)
```{r}



A6A9 <- ggplot(rtwl_data, aes(x = COL, y = Income)) + geom_point() + geom_point(aes(color = factor(RTWL)))
A6A9

A6A10 <- ggplot(rtwl_data, aes(x = URate, y = Taxes)) + geom_point() + geom_point(aes(color = factor(RTWL)))
A6A10

A6A10 <- ggplot(rtwl_data, aes(x = URate, y = Taxes)) + geom_point() + geom_point(aes(color = factor(RTWL)))
A6A10

A6A11 <- ggplot(rtwl_data, aes(x = COL, y = PD)) + geom_point() + geom_point(aes(color = factor(RTWL)))
A6A11
```

