---
title: "Regression analysis on death rate data"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---


```{r}

library(dplyr)
library(scatterplot3d)
library(psych)
library(ggplot2)
library(car)
library(gclus)
library(GGally)
library(fields)

deathrate <- read.table("x28_NoHeader", header = FALSE)

colnames(deathrate) <- c("Index",
                         "A1",
                         "A2",
                         "A3",
                         "A4",
                         "A5",
                         "A6",
                         "A7",
                         "A8",
                         "A9",
                         "A10",
                         "A11",
                         "A12",
                         "A13",                         
                         "A14",
                         "A15",
                         "B")
# colnames(deathrate) <- c("Index",
#                          "percip",
#                          "Jantemp",
#                          "Julytemp",
#                          "popover65",
#                          "housesize",
#                          "schooling",
#                          "kitchen",
#                          "urbanpop",
#                          "nonwhite",
#                          "officeworkers",
#                          "poorfamily",
#                          "hydrocarbonpollution",
#                          "nitrogenpollution",                         
#                          "sulfardioxidepollution",
#                          "humidity",
#                          "Death Rates")

# 
# Index
# A1 average annual precipitation in inches
# A2 average January temperature in degrees Fahrenheit
# A3 average July temperature in degrees Fahrenheit
# A4 percent of 1960 SMSA population 65 years old or older
## A5 household size, 1960
## A6 schooling for persons over 22
# A7 household with full kitchens
## A8 population per square mile in urbanized areas
## A9 percent nonwhite population
# A10 percent office workers
## A11 poor families (annual income under $3000)
# A12 relative pollution potential of hydrocarbons
# A13 relative pollution potential of oxides of Nitrogen
# A14 relative pollution of Sulfur Dioxide
# A15 percent relative humidity, annual average at 1pm.
# B death rate
```


Here we will create a subset of the original dataframe with the desired features.
```{r}
myvars <- c("A5", "A6", "A9", "A11", "B")
deathrate_1 <- deathrate[myvars]
```

```{r}
 #Making a Basic scatter plot matrix
 
 scatterplotMatrix(~B+A5+A6+A9+A11,data=deathrate_1, smooth = FALSE, ellipse=
 FALSE, main="Simple Scatterplot Matrix")

```



```{r}
dta.r <- abs(cor(deathrate_1)) # get correlations
dta.col <- dmat.color(dta.r) # get colors
# reorder variables so those with highest correlation
# are closest to the diagonal
dta.o <- order.single(dta.r) 
cpairs(deathrate_1, dta.o, panel.colors=dta.col, gap=.5,
main="Variables Ordered and Colored by Correlation" )

dta.r
```








```{r}
# 3D Scatterplot with Coloring and Vertical Lines
# and Regression Plane 
 
attach(deathrate_1) 
s3d <-scatterplot3d(A6,A9,B, pch=16, highlight.3d=TRUE,
  type="h", main="3D Scatterplot")
fit <- lm(B ~ A6+ A9, data = deathrate_1) 
s3d$plane3d(fit)

A6A9 <- ggplot(deathrate_1, aes(x = A6, y = A9)) + geom_point() + geom_point(aes(color = B)) 
A6A9

attach(deathrate_1) 
s3d <-scatterplot3d(A5,A6,B, pch=16, highlight.3d=TRUE,
  type="h", main="3D Scatterplot")
fit <- lm(B ~ A5+ A6, data = deathrate_1) 
s3d$plane3d(fit)

A5A6 <- ggplot(deathrate_1, aes(x = A5, y = A6)) + geom_point() + geom_point(aes(color = B)) 
A5A6

attach(deathrate_1) 
s3d <-scatterplot3d(A6,A11,B, pch=16, highlight.3d=TRUE,
  type="h", main="3D Scatterplot")
fit <- lm(B ~ A6+ A11, data = deathrate_1) 
s3d$plane3d(fit)

A6A11 <- ggplot(deathrate_1, aes(x = A6, y = A11)) + geom_point() + geom_point(aes(color = B)) 
A6A11
```



```{r}
#Finding correlation coeff r with p-values
#install.packages("psych")
library(psych)
corr.test(deathrate_1, y = NULL, use =
"pairwise",method="pearson",adjust="holm",alpha=.05)
```





# ```{r}
# ggcorr(deathrate_1, palette = "RdYlGn", name = "rho", 
#        label = FALSE, label_color = "black")
# 
# # Correlation plot
# ggcorr(deathrate_1, palette = "RdBu", label = TRUE)
# ```

```{r}
ggcorr(deathrate_1, palette = "RdYlGn", name = "rho", 
       label = FALSE, label_color = "black")

# Correlation plot
ggcorr(deathrate_1, palette = "RdBu", label = TRUE)
```

```{r}
ggpairs(deathrate_1, columns = 1:ncol(deathrate_1), title = "",  
  axisLabels = "show", columnLabels = colnames(deathrate_1[, 1:5]))
```
## A5 household size, 1960
## A6 schooling for persons over 22
# A7 household with full kitchens
## A8 population per square mile in urbanized areas
## A9 percent nonwhite population
# A10 percent office workers
## A11 poor families (annual income under $3000)
The Model we will choose is 

```{r}
ggpairs(deathrate_1, columns = 2:ncol(deathrate_1), title = "",  
  axisLabels = "show", columnLabels = colnames(deathrate_1[, 2:5]))
```

** Y^i = A5X1 + A6X2 + A9X3 + A11X4 + error^i, error is normal rv with mean=o and constant varaince for all i. **

```{r}
fit1 <- lm(formula = deathrate_1$B ~ deathrate$A5 + deathrate_1$A6 + deathrate_1$A9 + deathrate_1$A11)

fit2 <- lm(formula = deathrate_1$B ~deathrate_1$A6 + deathrate_1$A9 + deathrate_1$A11)

summary(fit1)
anova(fit1, fit2)
summary(fit2)
```


```{r}
shapiro.test(fit2$residuals)
```

```{r}
layout(matrix(c(1,2,3,4),2,2)) 
plot(fit2)
```
```{r}
hist(fit2$residuals)
```



```{r}
"A6"
summary(deathrate_1$A6)
"A9"
summary(deathrate_1$A9)
"A11"
summary(deathrate_1$A11)

```

```{r}
results = lm(B ~ A6 + A9 + A11, data=deathrate_1)

"Prediction: If the minimum value for all the predictors occured"
newdat1 = data.frame(A6 = 9, A9=0.8, A11 = 9.4)
predict(results, newdata = newdat1, interval="confidence")

"Prediction: If the mean of all the predictors occured"
newdat2 = data.frame(A6 = 10.97, A9=11.87, A11 = 14.37)
predict(results, newdata = newdat2, interval="confidence")

"Prediction: If the max of all the predictors occured"
newdat3 = data.frame(A6 = 12.3, A9=38.50, A11 = 26.40)
predict(results, newdata = newdat3, interval="confidence")

"Prediction: max min min"
newdat4 = data.frame(A6 = 12.3, A9=0.8, A11 = 9.4)
predict(results, newdata = newdat4, interval="confidence")
```

```{r}
summary(results)
```
Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output

Attaching package: ‘dplyr’

The following object is masked from ‘package:GGally’:

    nasa

The following objects are masked from ‘package:stats’:

    filter, lag

The following objects are masked from ‘package:base’:

    intersect, setdiff, setequal, union


Attaching package: ‘psych’

The following objects are masked from ‘package:ggplot2’:

    %+%, alpha

Loading required package: carData

Attaching package: ‘car’

The following object is masked from ‘package:psych’:

    logit

The following object is masked from ‘package:dplyr’:

    recode

Loading required package: cluster
Show in New WindowClear OutputExpand/Collapse Output

           A5        A6        A9       A11         B
A5  1.0000000 0.3950745 0.4188986 0.2599043 0.3573149
A6  0.3950745 1.0000000 0.2086791 0.4033385 0.5109837
A9  0.4188986 0.2086791 1.0000000 0.7053826 0.6437730
A11 0.2599043 0.4033385 0.7053826 1.0000000 0.4104873
B   0.3573149 0.5109837 0.6437730 0.4104873 1.0000000
R Console

Show in New WindowClear OutputExpand/Collapse Output


The following objects are masked from deathrate_1 (pos = 3):

    A11, A5, A6, A9, B

The following objects are masked from deathrate_1 (pos = 3):

    A11, A5, A6, A9, B

The following objects are masked from deathrate_1 (pos = 4):

    A11, A5, A6, A9, B

R Console





Show in New WindowClear OutputExpand/Collapse Output
Call:corr.test(x = deathrate_1, y = NULL, use = "pairwise", method = "pearson", 
    adjust = "holm", alpha = 0.05)
Correlation matrix 
       A5    A6    A9   A11     B
A5   1.00 -0.40  0.42  0.26  0.36
A6  -0.40  1.00 -0.21 -0.40 -0.51
A9   0.42 -0.21  1.00  0.71  0.64
A11  0.26 -0.40  0.71  1.00  0.41
B    0.36 -0.51  0.64  0.41  1.00
Sample Size 
[1] 60
Probability values (Entries above the diagonal are adjusted for multiple tests.) 
      A5   A6   A9  A11    B
A5  0.00 0.01 0.01 0.09 0.02
A6  0.00 0.00 0.11 0.01 0.00
A9  0.00 0.11 0.00 0.00 0.00
A11 0.04 0.00 0.00 0.00 0.01
B   0.01 0.00 0.00 0.00 0.00

 To see confidence intervals of the correlations, print with the short=FALSE option
Modify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output



Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output

Call:
lm(formula = deathrate_1$B ~ deathrate$A5 + deathrate_1$A6 + 
    deathrate_1$A9 + deathrate_1$A11)

Residuals:
     Min       1Q   Median       3Q      Max 
-108.201  -24.577   -0.139   23.442  100.826 

Coefficients:
                 Estimate Std. Error t value Pr(>|t|)    
(Intercept)     1530.0886   202.3078   7.563 4.51e-10 ***
deathrate$A5     -48.2837    45.9965  -1.050  0.29844    
deathrate_1$A6   -38.5904     7.3424  -5.256 2.47e-06 ***
deathrate_1$A9     5.8119     0.9036   6.432 3.20e-08 ***
deathrate_1$A11   -5.4068     1.9339  -2.796  0.00712 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 39.79 on 55 degrees of freedom
Multiple R-squared:  0.6185,	Adjusted R-squared:  0.5908 
F-statistic: 22.29 on 4 and 55 DF,  p-value: 5.57e-11

Analysis of Variance Table

Model 1: deathrate_1$B ~ deathrate$A5 + deathrate_1$A6 + deathrate_1$A9 + 
    deathrate_1$A11
Model 2: deathrate_1$B ~ deathrate_1$A6 + deathrate_1$A9 + deathrate_1$A11
  Res.Df   RSS Df Sum of Sq      F Pr(>F)
1     55 87094                           
2     56 88839 -1   -1744.9 1.1019 0.2984

Call:
lm(formula = deathrate_1$B ~ deathrate_1$A6 + deathrate_1$A9 + 
    deathrate_1$A11)

Residuals:
     Min       1Q   Median       3Q      Max 
-111.810  -23.740    0.083   21.385  103.480 

Coefficients:
                 Estimate Std. Error t value Pr(>|t|)    
(Intercept)     1337.5455    85.4275  15.657  < 2e-16 ***
deathrate_1$A6   -35.5427     6.7502  -5.265 2.30e-06 ***
deathrate_1$A9     5.4250     0.8258   6.569 1.78e-08 ***
deathrate_1$A11   -4.9799     1.8924  -2.632    0.011 *  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 39.83 on 56 degrees of freedom
Multiple R-squared:  0.6109,	Adjusted R-squared:   0.59 
F-statistic:  29.3 on 3 and 56 DF,  p-value: 1.591e-11

Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output
                                                                        
R Console


Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output

	Shapiro-Wilk normality test

data:  fit2$residuals
W = 0.99141, p-value = 0.9498

Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output

Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output





Show in New WindowClear OutputExpand/Collapse Output
[1] "A6"
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   9.00   10.40   11.05   10.97   11.50   12.30 
[1] "A9"
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   0.80    4.95   10.40   11.87   15.65   38.50 
[1] "A11"
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   9.40   12.00   13.20   14.37   15.15   26.40 
   
   
   
   
   
   
   
   
   
```{r}
# Below: We will make the REGRESSION "CUBE" for the three variables, and the predicted death rate, given # by a color gradient.

## A6 schooling for persons over 22
# A7 household with full kitchens
## A8 population per square mile in urbanized areas
## A9 percent nonwhite population
# A10 percent office workers
## A11 poor families (annual income under $3000)

level1 <- matrix(nrow = 18, ncol = 38)

x= seq(from=9.4, to=26.4, by = 1)
y= seq(.8,38.5,1)

for (j in 1:38){
    for (i in 1:18){
    level1[i,j] <- predict(results, newdata = data.frame(A6 = 9, A9=y[j], A11 = x[i]))
  
  }
}

# A6 = 10.40
level2 <- matrix(nrow = 18, ncol = 38)

x= seq(from=9.4, to=26.4, by = 1)
y= seq(.8,38.5,1)

for (j in 1:38){
    for (i in 1:18){
    level2[i,j] <- predict(results, newdata = data.frame(A6 = 10.40, A9=y[j], A11 = x[i]))
  
  }
}

# A6 = 11.05
level3 <- matrix(nrow = 18, ncol = 38)

x= seq(from=9.4, to=26.4, by = 1)
y= seq(.8,38.5,1)

for (j in 1:38){
    for (i in 1:18){
    level3[i,j] <- predict(results, newdata = data.frame(A6 = 11.05, A9=y[j], A11 = x[i]))
  
  }
}

# A6 = 11.50
level4 <- matrix(nrow = 18, ncol = 38)

x= seq(from=9.4, to=26.4, by = 1)
y= seq(.8,38.5,1)

for (j in 1:38){
    for (i in 1:18){
    level4[i,j] <- predict(results, newdata = data.frame(A6 = 11.50, A9=y[j], A11 = x[i]))
  
  }
}

# A6 = 12.30
level5 <- matrix(nrow = 18, ncol = 38)

x= seq(from=9.4, to=26.4, by = 1)
y= seq(.8,38.5,1)

for (j in 1:38){
    for (i in 1:18){
    level5[i,j] <- predict(results, newdata = data.frame(A6 = 12.30, A9=y[j], A11 = x[i]))
  
  }
}
```



```{r}
par(oma=c( 0,0,0,4)) # margin of 4 spaces width at right hand side
set.panel( 2,3) # 2X2 matrix of plots
image.plot(level1, zlim = c(773,1176))
title("A6=9")
image.plot(level2, zlim = c(773,1176))
title("A6=10.40")
image.plot(level3, zlim = c(773,1176))
title("A6=11.05")
image.plot(level4, zlim = c(773,1176))
title("A6=11.50")
image.plot(level5, zlim = c(773,1176))
title("A6=12.30")
```

```{r}
summary(fit2$B)

```


